%%
%% Template intro.tex
%%

\chapter{Introduction}
\label{cha:intro}

This thesis provides a detailed account of how an optimisation technique known
as Term Indexing was added to the \beagle\ first order logic theorem prover. This
introductory chapter provides an overview of the goals and motivation for this project,
along with a brief summary of what the whole paper will cover.

\section{Goals and Motivation}
\label{sec:mot}

The primary goal of this project is the implementation of Term Indexing
for the \beagle\ theorem prover. \Beagle\ implements a rather unique set of logic rules known as the \emph{\HSWAC} \cite{baum13}.
This logical calculus allows the inclusion of \emph{background reasoning} modules,
which implement a system of prior knowledge; for example the rules of integer arithmetic.

\Beagle\ was created rapidly as a demonstration and testbed for the capabilities
of its hierarchic reasoning system. As such, it is lacking many key features required
to make it a viable competitor against existing theorem provers. The most significant
feature which is yet to be implemented is \emph{Term Indexing}. Term Indexing exists
to efficiently locate logic expressions which are likely applicable to calculus \emph{inference rules},
used to generate new knowledge. Without this feature \beagle\ locates inference clauses
via a worst-case $O(n^2)$ search; trying each possible pair of clauses until a rule applies.

The specific indexing technique which will be implemented is Fingerprint Indexing;
a very recent technique developed by Stehpan \citeN{shulz12}. It works by computing
a set of features for a each term, known as their \emph{fingerprints}, which can then
be compared for compatibility. The details of this implementation form the core
of this thesis; which will include a detailed description of Fingerprint Indexing
and an account of how it was built and tested.
After covering the indexing implementation, this report will provide results measuring
how it has affected \beagle's performance. This will require test runs over a set
of logic problems for a variety of \beagle\ configurations.

\subsection{Cross Validating Fingerprint Indexing}

Fingerprint Indexing itself was introduced for the first
time in \cite{shulz12}; and at the time of writing the only
implementation of Fingerprint Indexing was for Shulz's tests. By implementing this new technique in \beagle\ 
we will be providing the Term Indexing field with more evidence regarding the
overall performance of Fingerprint Indexing.

Performance results for Fingerprint Indexing in \cite{shulz12} appear very promising;
and if the \beagle\ implementation can achieve similar improvements we will
have solidified Fingerprint Indexing as a major competitor in the field of Term Indexing.

\subsection{Improving Fingerprint Indexing}

Another side-goal of the project is to investigate and implement original
improvements to Fingerprint Indexing.
This can be achieved by \emph{tailoring} Fingerprint Indexing to \beagle\ 
and hierarchic reasoning. Specifically it may be possible to take
advantage of the large number of conditions that the \HSWAC\ places on
its inference rules.

It may also be possible to find improvements to Fingerprint Indexing which
apply more generally. General improvements will be explored; but considering
that tailored improvements would result in more significant performance increases they
are considered somewhat `out of scope' for this project.

\section{Structure of this Report}
\label{sec:framework}

%In this report we will...

\subsection{Background}

The background chapter (Chapter \ref{cha:background}) provides necessary background material required to understand
the concepts used in implementing Fingerprint Indexing. Obviously this includes
an in-depth explanation of precisely what Fingerprint Indexing is and how it works.
The chapter will also provide descriptions of several other Term Indexing techniques and
indicate some key advantages of Fingerprint Indexing.

Other background material includes a reference glossary of logic terminology and
a detailed specification of \beagle's \HSWAC.

\subsection{Method}

The method chapter (Chapter \ref{cha:method}) will cover in detail how Fingerprint
Indexing was added to \beagle. This includes creating the indexer itself, applying
it to \beagle's inference rules and investigating any original improvements.

Creating the the Fingerprint Index requires analysing \beagle's existing code
in order to understand how it implements logic expressions. We will then require
an implementation of any data structures required for Fingerprint Indexing along
with functions for adding to and retrieving from the index.

With the indexer built we will then investigate how it applies to \beagle's inference
rules. This will require the referring to the \HSWAC\ specification and another analysis of \beagle's existing code.
Only once all applicable areas of \beagle\ have been fully implemented with Fingerprint
Indexing will we consider some original improvements; such as tailoring indexing
to \beagle's peculiar calculus.

\subsection{Results}

The results chapter (Chapter \ref{cha:results}) provides test details
in order to confirm the effectiveness of Fingerprint Indexing. This will require
the selection of a set of logic problems to test against.
The chapter will provide comparisons for indexed and unindexed \beagle\ together
with comparisons between various configurations of Fingerprint Indexing.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "thesis"
%%% End: 
